---
title: "hw05"
author: "Qiaoyue Tang"
date: "2017/10/18"
output: github_document
---

```{r, include=FALSE}
library(gapminder)
library(dplyr)
library(ggplot2)
library(forcats)
```


### Factor Management

dataset: Gapminder

#### Drop Oceania

```{r}
dat1 <- gapminder %>% 
  filter(continent != c("Oceania")) %>% 
  droplevels()

t.droplevel <- data.frame(nrow(gapminder), nrow(dat1), nlevels(gapminder$continent), nlevels(dat1$continent))
names(t.droplevel) <- c("# rows before drop", "# rows after drop", "# levels before drop", "# levels after drop")
knitr::kable(t.droplevel)

gapminder %>% 
  filter(continent == "Oceania") %>% 
  nrow()
```

The gapminder dataset originally have 1704 rows. After drop, the number of rows decrease to 1680, where we checked that there are 24 observations with continent of "Oceania". The numbers match up. After we drop the level of continents, the number of levels change from 5 to 4. The other variables are not affected.

#### Reorder Levels

```{r}
fct_reorder(dat1$continent, dat1$gdpPercap, fun = max, desc = F) %>% 
  levels()
```

The levels of continent are reordered so that the maximum of GDP are in ascending order. We can check if it is indeed in this order by:

```{r}
dat1 %>% 
  group_by(continent) %>% 
  summarise(max(gdpPercap))
```

So the order should be Africa, Americas, Europe then Asia which is the same as ordered by the previous function.

```{r}
fct_reorder(dat1$country, dat1$lifeExp, fun = mean, desc = T) %>% 
  levels() %>% 
  head()
```

The levels of country are reordered so that the mean of life expectancy are in decending order.
Check this by:

```{r}
check<- dat1 %>% 
  group_by(country) %>% 
  summarise(mean(lifeExp)) %>% 
  data.frame()
names(check) <- c("country", "mean_lifeExp")
check[with(check, order(mean_lifeExp)), ] %>% 
  head()
```

We can see that the new order is exactly ordered by desending mean of life expectancy. (Here we only look at head to avoid lengthy dataframes.)

#### Common Part

Explore the effects of **arrange** 

```{r}
gapminder_subset<- gapminder %>% 
  filter(continent %in% c("Asia"))

dat2 <- gapminder %>%
  filter(continent %in% c("Asia")) %>% 
  arrange(lifeExp)

head(levels(dat2$country), n=10)
head(levels(gapminder_subset$country), n=10)

ggplot(dat2, aes(x=lifeExp, y=country)) + geom_line() + ggtitle("Plot after Arrange") + theme(plot.title = element_text(hjust = 0.5))

ggplot(gapminder_subset, aes(x=lifeExp, y=country)) + geom_line() + ggtitle("Plot before Arrange") + theme(plot.title = element_text(hjust = 0.5))
```

We can see that arrange does not have an effect on the plots, and the levels are the same before and after arrange.

Explore the effects of **reorder and arrange**

```{r}
dat2$country <- fct_reorder(dat2$country, dat2$lifeExp, fun = mean, desc = T)

head(levels(dat2$country), n=10)

ggplot(dat2, aes(x=lifeExp, y=country)) + geom_line() + ggtitle("Plot before Arrange") + theme(plot.title = element_text(hjust = 0.5))
```

We can see that the levels are now arranged in the order of decreasing mean of life expectancy, and the graph is changed as well. The plot is now using the reordered levels.