---
title: "hw03"
author: "Qiaoyue Tang"
date: "2017/9/30"
output: github_document
---

```{r setup, include=FALSE}
library(gapminder)
library(ggplot2)
library(dplyr)
```

### Task 1: Get the maximum and minimum of GDP per capita for all continents.

```{r}
gapminder %>%
  group_by(continent) %>% 
  summarize(max_gdp=max(gdpPercap), min_gdp=min(gdpPercap)) %>% 
  mutate(diff=max_gdp-min_gdp)
```
From the table we can see that the maximun of GDP per Capita over all appears in Asia, and the minimum appears in Africa. I've also included a difference column which tells the difference between the maximum GDP and the minimum GDP. We can see that the largest gap occurs in Asia.

One problem that I hope to solve is to make min_gdp also with 2 digits of decimals, but since summarize does not have an argument that sets decimal places.

### Task 2: Look at the spread of GDP per capita within the continents.

```{r}
gapminder %>%
  select(gdpPercap, continent) %>% 
  ggplot(aes(x=continent, y=gdpPercap)) + 
              geom_point(alpha=0.5) + 
              geom_boxplot() +
              ggtitle("Boxplot of GDP per Capital Across Continents") +
              theme(plot.title = element_text(hjust = 0.5)) +
              labs(x="Continent", y="GDP per Capita")
              
```
From the boxplot we can see that Asia has the largest variation of GDP per Capita. Even though the middle 50% data are not as variable as Europe, it has a lot more extreme values than the other continents. Oceania seems to have the smallest varaition.

One interesting issue: when I put labs() after theme(), the title alignment is not adjusted. Not sure why this is the case.

### Task 3: Compute a trimmed mean of life expectancy for different years.

```{r}
dat1 <- gapminder %>%
  select(country, lifeExp, year) %>% 
  group_by(country) %>% 
  summarise(weighted_lifeExp = weighted.mean(lifeExp, c(1:12)/78))

dat2 <- gapminder %>%
  select(country, lifeExp, year) %>% 
  group_by(country) %>%
  summarize(mean(lifeExp))

inner_join(dat1, dat2, by="country")
```
dat1 provides the trimmed life expectancy and dat2 provides the ordinary mean of life expectancy. I joined them together in one table using inner_join. The trimmed mean weights are assigned as 1/sum(1:12), 2/sum(1:12), etc. which puts larger weight on later years and smaller weight on previous years. It would make sense in the case where we believe data collection process are getting more precise, so we want the most accurate data to reflect the true mean.
